---
title: "Introduction to Scientific Computing for Biologists"
subtitle: 'ISCB20.09 - Introduction to R'
author: | 
  | Md. Jubayer Hossain
  | https://jhossain.me/
  | jubayer@hdrobd.org 
institute: | 
  | Founder 
  | Health Data Research Organization
  | Lead Organizer
  | Scientific Computing for Biologists 
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  beamer_presentation:
    highlight: espresso
    fonttheme: 'structurebold'
classoption: 'aspectratio=169'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA)
```

# Section-4: Data Types in R 
## Data Types in R{.t}
R has five basic data types 

- character(e.g: 'abul', "abul")
- numeric(e.g: 2, 3)
- integer(e.g: 5L)
- complex(e.g: 5i)
- logical(True/False)

## Numbers{.t}
* Numbers in R generally treated as numeric objects(i.e. double precision real numbers)
* If you explicitly want an integer, you need to specify the `L` suffix.
* There is also a special number `Inf` which is represents infinity; e.g. `1 / 0`
* `Inf` can be used in ordinary calculations; e.g. `1/Inf` is `0`
* The value `NaN` represents an undefined value("not a number"); e.g. `0/0` 
* `NaN` can also be thought of as a missing value. 

The numeric constants are 

- integer(`L`)
- double
- complex(`i`) 

## Decimal vs Double vs Float{.t}
The Decimal, Double, and Float variable types are different in the way that they store the values. Precision is the main difference where float is a single precision (32 bit) floating point data type, double is a double precision (64 bit) floating point data type and decimal is a 128-bit floating point data type.

* Float - 32 bit (7 digits)
* Double - 64 bit (15-16 digits)
* Decimal - 128 bit (28-29 significant digits)

## Characters{.t}
Character constants can be represented using either single quotes ('') or double quotes ("") as delimiters.

## Atributes{.t} 
R objects can have attributes 

* names, dimnames 
* dimensions(e.g. matrices, arrays)
* class 
* length 
* other user-defined attributes/metadata

Attributes of an object can be accessed using the `attributes()` function.

## Built-in Constants in R{.t}

```{r}
pi # value of pi 
LETTERS
```

```{r}
month.name 
```



