---
title: "Introduction to Scientific Computing for Biologists"
subtitle: 'ISCB20.09 - Introduction to R'
author: | 
  | Md. Jubayer Hossain
  | https://jhossain.me/
  | jubayer@hdrobd.org 
institute: | 
  | Founder 
  | Health Data Research Organization
  | Lead Organizer
  | Scientific Computing for Biologists 
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  beamer_presentation:
    highlight: espresso
    fonttheme: 'structurebold'
classoption: 'aspectratio=169'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA)
```
# Section-5: Operators

## Operators in R{.t}
- R has many operators to carry out different mathematical and logical operations.
- Operators in R can mainly be classified into the following categories. 
  1. Arithmetic Operators 
  2. Relational Operators
  3. Logical Operators 
  4. Assignment Operators 

## Arithmetic Operators{.t}
These operators are used to carry out mathematical operations like addition and multiplication. The arithmetic operators are addition(+), subtraction(-), multiplication(*), division(/),  exponent(^), modulus(%%) and integer division(%/%). 

|Operators | Description|
|----------|:-----------|
|+|Addition|
|-|Subtraction|
|*|Multiplication| 
|/|Division | 
|%%|Modulus |
|%/%|Integer Division |



```{r}
x <- 10 
y <- 2 
```


```{r}
# Addition
x+y
# Subtraction
2-5
```


## Arithmetic Operators(Cont...){.t}
```{r}
# Multiplication
2 * 5 

# Division
2 / 5 

# Exponent 
2 ^ 5
```

## Arithmetic Operators(Cont...){.t}

```{r}
# Modulus(Remainder from division)
2 %% 5

# Integer Division 
2 %/% 5
```

## Bolean Operators{.t}
```{r}
x <- TRUE 
y <- TRUE 
a <- T 
b <- T 
```



## Relational Operators{.t}
```{r}
x <- 10
y <- 5 
# Equality 
x == y 

# Inequality
x != y 

# Less Than 
x < y 
```

## Logical Operators{.t}
```{r}
# Logical NOT(!)
! TRUE
! FALSE   
```
## Logical Operators(Cont...){.t}
```{r}
# Logical AND(&&)
TRUE && TRUE
TRUE && FALSE 
FALSE && FALSE 
```

## Logical Operators(Cont...){.t}
```{r}
# Logical OR(|)
TRUE | TRUE 
TRUE | FALSE 
FALSE | FALSE    
```


## Relational Operators(Cont...){.t}
```{r}
# Greater Than 
x > y 
# Less or Equal 
x <= y 
# Greater or Equal
x >= y 
```