---
title: "Introduction to Scientific Computing for Biologists"
subtitle: 'ISCB20.09 - Introduction to R'
author: | 
  | Md. Jubayer Hossain
  | https://jhossain.me/
  | jubayer@hdrobd.org 
institute: | 
  | Founder 
  | Health Data Research Organization
  | Lead Organizer
  | Scientific Computing for Biologists 
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  beamer_presentation:
    highlight: espresso
    fonttheme: 'structurebold'
classoption: 'aspectratio=169'
---


# Section-3: Variables and Reserved Keywords

## Variables(Identifiers) in R{.t}
- Variables are used to **store data**, whose value can be changed according to our need. \pause 
- A variable is a name given to a memory location, which is used to store values in a computer program. \pause 
- Variables in R programming can be used to store numbers (real and complex), words, matrices, and even tables. \pause 
- R is a dynamically programmed language which means that unlike other programming languages, we do not have to declare the data type of a variable before we can use it in our program. \pause 
- Unique name given to variable (function and objects as well) is **identifier**.


## Rules for writing Identifiers in R{.t}
- Identifiers can be a combination of letters, digits, period (.) and underscore (_). \pause 
- It must start with a letter or a period. If it starts with a period, it cannot be followed by a digit.\pause 
- It should not start with a number (e.g:- 2x) \pause 
- It should not start with a dot followed by a number (e.g:- .2x) \pause 
- It should not start with an underscore (e.g:- _x) \pause 
- Reserved words in R cannot be used as identifiers(e.g: TRUE, FALSE) \pause 

Basically, there are 5 naming conventions 

- alllowercase: e.g. `myname`
- period.separated: e.g. `new.name`
- underscore_separated: e.g. `my_name`
- lowerCamelCase: e.g. `myName`
- UpperCamelCase: e.g. `MyName`

## Creating Variables{.t}
Using equal(`=`) operator
```{r echo=TRUE}
x = 10 
```
Using leftward(`<-`) operator 
```{r echo=TRUE}
y <- 15  
```

Using rightward(`->`) operator 
```{r echo=TRUE}
z <- 20 
```

## Reserved Keaywords in R{.t}
- Don’t use any reserved keyword as variable name. List all of reserved words in R by using (?Reserved). 
- The reserved words in R’s parser are - if else repeat while function for in next break - TRUE FALSE NULL Inf NaN NA NA_integer_ NA_real_ NA_complex_ NA_character_


## Entering Input{.t}
At the R prompt/console we type expressions. 
```{r echo=TRUE}
num <- 10
```
The `<-` symbol is the **assignment** operator. \
The grammar of the language determines whether an expression is complete or not.

## Evaluation{.t}
When a complete expression is entered at the R console, it is evaluated and the result of evaluated expression is returned. The result may be auto-printed. 
```{r}
x <- 10
x
```

```{r}
x <- 10
print(x)
```

```{r}
x <- 10
cat(x)
```
  
## Comments in R{.t}
The # character indicates a comment. 
```{r}
x <- 10 # This is a comment
```
Anything to the right of the # (inlcuding the # itself) is ignored. \
This is only comment character in R. \
R does not support multi-line comments or comment block.

